---
title: "Practice with t-tests"
output: html_document
date: "2022-10-27"
---

## Set up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

Load in the data:
```{r}
bf = read.csv(here("Data", "body_fat_percentage.csv"))
```

Gender as a factor:
```{r}
bf$gender = as.factor(bf$gender)
```

## Reviewing the t-test

Let's conduct a 2-sample t-test to compare body fat percentage between men and women.

The data set presents the body fat percentage for a sample of men and women who all work out at the same gym. In general, men are leaner than women, with a body fat percentage centering around 15% whereas for women it tends to center around 20%. Let’s see if the same is true of these data.

## Instructions

Use the information you learned from reading chapter 5 and doing homework 21 to conduct a t-test comparing mean body fat percentage between the male and female groups. But don’t just rush to the test – follow the recommended analysis workflow:

1.  Plot the data and look at expected relationships

```{r}
ggplot(bf, aes(x = gender, y = body_fat_percent))+
  geom_boxplot(outlier.color = "red", color = c("gray", "black"))+
  geom_point(size = 5, color = "blue", alpha = .5)+
  theme_bw()
```

2.  Calculate the mean and standard deviation for the data and examine the resulting output.

```{r}
bf |> group_by(gender) |> summarise(
  meanbf = mean(body_fat_percent),
  sdbf = sd(body_fat_percent))
```

Males have a lower mean body fat percentage and are slightly more variable than females.

3.  Discuss with your partner:
    -   Review in the book: What are the assumptions about the data appropriate for a t-test? When you examine the plots and summary data, do they appear to meet those assumptions?
    
    
The equal variance assumption could be violated because they are not equal (thought they are close,) but the independence condition should be met by these data. We are not sure about normality, but the plots look approximately normal.
    
   -   Based on examination of the plots and summary data, do you expect that there is a statistically significant difference in mean body fat percentage between males and females in this sample?
4.  Conduct a t-test and then interpret the results. What do they tell you about the null vs. alternative hypotheses?

```{r}
t.test(body_fat_percent~gender, data = bf)
```

Reject the null. We have statistically discernible evidence that there is a difference in mean body fat percentage between male and female groups. Men have discernibly less body fat than women (t = 2.86, df = 20.96, p = 0.00932)

